\bf Baseline Drift Estimation for Time Series Data Using Quantile Trend Filtering

### 2019-02-08 Manuscript/detrendify_JCGS.tex
Eric's edits.

\documentclass[12pt]{article}
\RequirePackage{amsthm,amsmath,amsbsy,amsfonts}
\usepackage{bm,graphicx,psfrag,epsf}
%\usepackage{enumerate}
\usepackage{natbib}
\usepackage{url} % not crucial - just used below for the URL 
\usepackage{placeins}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{subfig} %% For \subfloat command

@@

%% Math notation
\newcommand{\Tra}{^{\sf T}} % Transpose
\newcommand{\Inv}{^{-1}} % Inverse
\newcommand{\tr}{\operatorname{tr}} % Trace
\newcommand{\Ker}{\operatorname{Ker}} % Kernel
\def\vec{\mathop{\rm vec}\nolimits}
\newcommand{\amp}{\mathop{\:\:\,}\nolimits}
\newcommand{\bic}{\text{BIC}}

@@

	\bigskip
	\begin{abstract}
		We address the problem of estimating smoothly varying baseline trends in time series data using quantile trend filtering. We extend the basic framework to ensure non-crossing while estimating multiple trends simultaneously and implement an alternating direction method of moments (ADMM) algorithm for series that cannot be processed simultaneously. Timing experiments demonstrate the usefulness of our ADMM algorithm even in cases where the series can be processed simultaneously. We also address smoothing parameter selection and propose a modified criteria based on the extended Bayesian Information Criteria. We demonstrate both through simulation studies and an application to low cost air quality sensor data that our model provides better or comparable estimates than existing methods of non-parametric quantile trends and improves signal classification of low-cost air quality sensor output. 
	\end{abstract}

@@

	In applications spanning the fields of chemistry \citep{Ning2014}, macroeconomics \citep{yamada2017estimating}, environmental science \citep{brantley2014mobile}, and medical sciences \citep{pettersson2013algorithm, marandi2015qualitative}, scalar time series are observed and assumed to consist of a slowly varying trend and other more rapidly varying components. \cite{Kim2009} addressed the case when the series consisted of just a trend and a random noise component and proposed the use of \textit{$\ell_1$-trend filtering} to estimate trends that piecewise linear or piecewise polynomial. \cite{Tib2014} later demonstrated that empirically the trend filtering estimates adapt to the local level of smoothness better than the more common smoothing splines \Halley{Eric}{If he proved theorems, minimax stuff, we should state that.}. In the trend filtering problem, given observations $y(t)$ with $t=1,\ldots,n$, the trend, $\theta(t)$, is estimated by solving the following convex optimization problem.
	\begin{eqnarray}
	\underset{\V{\theta}}{\text{minimize}}\; \frac{1}{2} \lVert \V{y} - \V{\theta} \rVert_2^2 + \lambda \lVert \Mn{D}{k+1}\V{\theta} \rVert_1,
	\end{eqnarray}
	where $\lambda \geq 0$ is a regularization parameter and the matrix $\Mn{D}{k+1} \in \Real^{(n - k -1) \times n}$ is the discrete difference operator of order $k+1$. To understand the purpose of penalizing $\Mn{D}{k+1}\V{\theta}$ consider the difference operator when $k = 0$.
	\begin{eqnarray}
	\Mn{D}{1} & = & \begin{pmatrix}
	-1 & 1 & 0 & \cdots & 0 & 0 \\
	0 & -1 & 1 & \cdots & 0 & 0 \\
	\vdots & & & & & \\
	0 & 0 & 0 & \cdots & -1 & 1 \\
	\end{pmatrix}.
	\end{eqnarray}
	Thus, $\lVert \Mn{D}{1}\V{\theta} \rVert_1 = \sum_{i=1}^{n-1} \lvert \theta_i - \theta_{i+1} \rvert$ which is just total variation denoising in one dimension. The penalty term incentivizes solutions which are piece-wise constant. For $k \geq 1$, the difference operator $\Mn{D}{k+1} \in \Real^{(n-k-1) \times n}$ is defined recursively as follows
	\begin{eqnarray}
	\Mn{D}{k+1} & = & \Mn{D}{1}\Mn{D}{k}.
	\end{eqnarray}
	By penalizing the $k+1$ fold composition of the discrete difference operator, we obtain solutions which are piecewise polynomials of order $k$.

	\Fig{tpn} shows that trend filtering provides excellent estimates of trends when the only components of the time series are the trend and random noise. In some applications, however, time series may also include a rapidly varying signal component in addition to the trend and noise. \Fig{tpspn} shows that in these cases the trend filtering estimate over-estimates the trend in the places where signal is present.

	\begin{figure}[t]
		\centering
		\subfloat[Slow Trend + Noise]{\label{fig:tpn}
		\includegraphics[width = 0.45\linewidth]{Figures/trend_filter_eg1.png}}
		\subfloat[Slow Trend + Spikes + Noise]{\label{fig:tpspn}
		\includegraphics[width = 0.45\linewidth]{Figures/trend_filter_eg2.png}}
		\caption{Two examples of trend filtering solutions. The true trend is shown in black, while the estimated trend is shown in red.}		
		\label{fig:trendfilter}
	\end{figure}

	One application in which the observed time series consists of a slowly varying baseline, non-negative signal, and rapidly varying noise is the output of low cost air quality sensors. The use of low-cost, portable, air quality sensors has increased dramatically in the last decade. These sensors can provide an un-calibrated measure of a variety of pollutants in near real time, but deriving meaningful information from sensor data remains a challenge \citep{snyder2013changing}. The ``SPod" is a low-cost sensor currently being investigated by researchers at the U.S. Environmental Protection Agency to detect volatile organic compound (VOC) emissions from industrial facilities \citep{thoma2016south}. To reduce cost and power consumption of the SPod, the temperature and relative humidity of the air presented to the photoionization detectors (PIDs) is not controlled and as a result the output signal exhibits a slowly varying baseline drift on the order of minutes to hours. \Fig{raw_spod} provides an example of 3 co-located SPod sensor outputs from near an industrial facility. While all of the sensors response to the pollutant signal, the amount of baseline drift varies from one node to another.

	\begin{figure}[b!]
		\caption{Example of 3 co-located SPod PID sensor readings.}
		\includegraphics[width = \linewidth]{Figures/uncorrected_data.png}
		\label{fig:raw_spod}
	\end{figure}

	In this application, as well as those described in \cite{Ning2014}, \cite{marandi2015qualitative}, and \cite{pettersson2013algorithm}, it is not the mean trend that is desired but rather the baseline trend, which can be thought of as a trend in a low quantile of the data.  \cite{Koenker1978} were the first to propose substituting the squared loss typically used in regression with the check loss function  \Eqn{checkloss} to estimate a conditional quantile instead of the conditional mean.

	\begin{equation}
	\label{eq:checkloss}
	 \rho_{\tau}(z) = \sum_{i=1}^n z_i(1-\mathbf{I}(z_i<0)).
	\end{equation}

	Later, \cite{KoenkerNgPortnoy1994} addressed the issue of estimated quantile curves by combining the check loss penalty with trend filtering penalty with $k = 2$ which produces quantile trends that are piecewise linear. Other approaches to estimating quantile trends have also been proposed. Rather than using the trend filtering $\ell_1$-norm to penalize the discrete differences, \cite{nychka1995nonparametric} used the smoothing spline penalty based on the $\ell_2$-norm solving the optimization problem in \Eqn{smoothingspline}.

	\begin{equation*}
	\label{eq:smoothingspline}
	\rho_{\tau}(y - f) + \lambda\int (f''(t))^2 dt, 
 	\end{equation*}

	where $f(t)$ is a smooth function of time and $\lambda$ is a tuning parameter that controls the degrees of smoothing. \cite{Oh2011} proposed an algorithm for solving this problem based on approximating the checkloss function with a similar differentiable function.

@@

	We propose to use the trend filtering penalty combined with the check loss function to produce quantile trends that are piecewise quadratic. The formulation was proposed by \cite{Kim2009} as a possible extension of $\ell_1$-trend filtering but not studied. Moreover we extend the basic framework to ensure non-crossing while modeling multiple quantiles. We also implement a parallel ADMM algorithm for series that are too large to be computed simultaneously and propose a modified criteria for choosing the smoothing parameter. We demonstrate through simulation studies that our proposed model provides better or comparable estimates of non-parametric quantile trends than existing methods and is a more effective method of drift removal for low-cost air quality sensors.

@@

	We combine the ideas of quantile regression and trend filtering. For a single quantile level $\tau$ the estimation of the quantile trend filtering model can be posed as the following optimization problem.
	\begin{eqnarray}
	\label{eq:quantile_trend}
	\underset{\theta}{\min}\; \rho_\tau(y - \theta) + \lambda \lVert \Mn{D}{k+1} \theta \rVert_1,
	\end{eqnarray}
	where $\lambda$ is a non-negative regularization parameter. We address the problem of choosing $\lambda$ in Section \ref{sec:lambda_choice}. As with classic quantile regression, the quantile trend filtering problem is a linear program which can be solved by a number of free or commercial solvers. In many cases, including ours, we are interested in estimating multiple quantiles simultaneously. We also want to ensure that our quantile estimates are valid by enforcing the constraint that if $\tau_2 > \tau_1$ then $Q(\tau_2) \ge Q(\tau_1)$. Given quantiles $\{\tau_1, \ldots, \tau_J\}$ such that $\tau_1 < \tau_2 < \ldots < \tau_J$, the optimization problem becomes

	\begin{eqnarray}
	\label{eq:noncross_trend}
	\underset{\theta_1, \ldots, \theta_J}{\min}\; \sum_{j=1}^J \left [\rho_{\tau_j}(y - \theta_{j}) + 
	\lambda_j \lVert \Mn{D}{k+1} \theta_j \rVert_1 \right ] \\
	 \text{subject to: }\; \theta_{1}(t) \le \theta_{2}(t) \le \ldots \le \theta_{J}(t) \text{ for all } t,
	\end{eqnarray}

	where $\theta_j \in \mathcal{R}^n$. The additional constraints are linear in the parameters so the non-crossing quantile trends can still be estimated by a number of available solvers. In the rest of this paper we rely on the commercial solver Gurobi \citep{gurobi} and its R package implementation, but it can easily be substituted for a free solver such as the Rglpk package by \cite{rglpk}. The number of parameters to be estimated in this problem is equal to the number of observations multiplied by the number of quantiles of interest. As the size of the data and the number of quantiles grows, all solvers will eventually break.

@@

	\begin{align*}
	y_1(t) = y(t) & \mbox{~~if~~} 1 \le t \le u_{1}\\
	y_2(t) = y(t) & \mbox{~~if~~} l_{2} \le t \le u_{2} \\
	\cdots & \\
	y_M(t) = y(t) & \mbox{~~if~~} l_{M} \le t \le  n\\
	\end{align*}

	 with boundaries $1 < l_{2} < u_{1} < l_{3} < u_{2} < l_{4} < u_{3} < \ldots< n$. An illustration is given in \Fig{windows}. Given quantiles $\tau_1 < \ldots < \tau_J$ to be estimated we define $\theta_{j,m}(t)$ as the value of the $\tau_j$\textsuperscript{th} quantile trend in window $M$ at time point $t$. In order to write out the constraint that the overlapping sections must be equal we define a consensus variable

	 \begin{align*}
		 \bar{\theta}_{j,m} =&  g(\theta_{j, m-1}, \theta_{j,m}, \theta_{j,m+1}) \\
		 =& \begin{cases} 
			 \frac{\theta_{j,m-1}(t)+\theta_{j,m}(t)}{2} & \mbox{if~~} l_{m} \le t \le u_{m-1}  \\
			 \theta_{j,m}(t) & \mbox{if~~} u_{m-1} \le t \le l_{m+1}  \\
			 \frac{\theta_{j,m}(t)+\theta_{j,m+1}(t)}{2} & \mbox{if~~} l_{m+1} \le t \le u_{m}  
			 \end{cases},
	\end{align*}
	defining $\theta_{j,M+1} = \theta_{j,M}$ and $\theta_{j,0} = \theta_{j,1}$. Our windowed quantile trend optimization problem can then be written as 
	 \begin{align*}
		 \label{eq:quantile_windows}
		 &\sum_{m=1}^M\underset{\theta_{1,m}, \ldots, \theta_{J,m}}{\min}\; \sum_{j=1}^J \left [\rho_{\tau_j}(y_m - \theta_{j,m}) + 
		 \lambda_j \lVert \Mn{D}{k+1} \theta_{j,m} \rVert_1 \right ] \\
		 &\text{subject to: }\; \\
		 &\qquad \theta_{1,m}(t) \le \theta_{2,m}(t) \le \ldots \le \theta_{J,m}(t) \text{ for all } m,t \\
		 &\qquad \theta_{j,m}(t) = \bar{\theta}_{j,m}(t) \text{ for all } j, m, t
	 \end{align*}
	 Defining Lagrange multipliers $\omega_{j,m}$, and penalty parameter $\gamma > 0$ we can write the augmented Lagrangian for finding the trends in window $m$:
	 \begin{equation*}
	 \mathcal{L}(\theta_{j,m}, \bar{\theta}_{j,m}, \omega_{j,m}) = \sum_{j=1}^J\rho_{\tau_j}(y_m - \theta_{j,m})+\lambda \lVert \Mn{D}{k+1}\theta_{j,m}\rVert_1 +  \omega_{j,m}^T(\theta_{j,m} - \bar{\theta}_{j,m}) + 
	 \frac{\gamma}{2}||\theta_{j,m} - \bar{\theta}_{j,m}||_2^2
	 \end{equation*}
	 We then estimate the trend separately in each window, which can be done in parallel, while constraining the overlapping pieces of the trends to be equal as outlined in algorithm 1.

@@

	\begin{algorithm}
		\caption{ADMM algorithm for quantile trend filtering with windows}\label{euclid}
		\begin{algorithmic}		
			\State Define $D = D^{(k+1)}$. 
			\State \textbf{initialize:} \\
				 $\theta_{j,m}^{(0)} = \arg \min \sum_{j=1}^J\rho_{\tau_j}(y_m - \theta_{j,m})+\lambda \lVert D\theta_{j,m}\rVert_1$ subject to $\theta_{1,m}(t) < \ldots<\theta_{J,m}(t)$ for all $t$. \\		
			 	 $\omega_{j,m}^{(0)} = \mathbf{0}$	
			\Repeat{}
			\State  		
			$\bar{\theta}_{j,m}^{(q)} = g(\theta_{j, m-1}^{(q-1)}, \theta_{j,m}^{(q-1)}, \theta_{j,m+1}^{(q-1)})$
			\State 
			$\omega_{j,m}^{(q)} = \omega_{j,m}^{(q-1)} + \gamma(\theta_{j,m}^{(q-1)} - \bar{\theta}_{j,m}^{(q)})$	
			\State
				$\theta_{j,m}^{(q)} = \arg\min \mathcal{L}(\theta_{j,m}, \bar{\theta}_{j,m}^{(q-1)}, \omega_{j,m}^{(q-1)})$			
			 subject to $\theta_{1,m}(t) < \ldots<\theta_{J,m}(t)$ for all $t$.
			\Until {convergence}
			\State \textbf{return} Non-overlapping sequence of $\bar{\theta}_{j,m}^{(q)}$ for all $j$, $m$.

@@

	We measure convergence use the stopping criteria described by \cite{boyd2011distributed}. The criteria are based on the primal and dual residuals which represent the residuals for the primal and dual feasibility, respectively. The primal residual, 
	\begin{equation}
	r_p^{(q)} = \sqrt{\sum_{m=1}^M\sum_{j=1}^J\lVert\theta_{j,m}^{(q)} - \bar{\theta}_{j,m}^{(q)}\rVert_2^2},
	\end{equation}
	represents the difference between the trend values in the windows and the consensus trend value while the dual residual,
	\begin{equation*}
	r_d^{(q)} = \gamma\sqrt{\sum_{m=1}^M \sum_{j=1}^J\lVert\bar{\theta}_{j,m}^{(q)} - \bar{\theta}_{j,m}^{(q-1)}\rVert_2^2},
	\end{equation*}
	represents the change in the consensus variable from one iterate to the next. The algorithm is stopped when 
	\begin{align}
		&r_p^{(q)} < \epsilon_{abs}\sqrt{nJ} + \epsilon_{rel}\underset{m}{\max}\left[\max 
		\left(\sqrt{\sum_{j=1}^J \lVert\theta_{j,m}^{(q)}\rVert_2^2}, \sqrt{\sum_{j=1}^J \lVert \bar{\theta}_{j,m}^{(q)} \rVert_2^2} \right )\right]\\
		&r_d^{(q)} < \epsilon_{abs}\sqrt{nJ} + \epsilon_{rel}\sqrt{\sum_{m=1}^M\sum_{j=1}^J\lVert \omega_{j,m}^{(q)}\rVert_2^2}
	\end{align}

	\begin{figure}
		\centering
		\caption{Trend fit with our ADMM algorithm with 3 windows which converged in 7 iterations compared to trend from simultaneous fit.}
		\includegraphics[width = 0.8\linewidth]{Figures/admm_windows.png}
	\end{figure}

	Timing experiments illustrate the advantages of using our ADMM algorithm even on datasets where solving the problem simultaneously is possible. For each data size, $n$, 25 datasets were simulated using the peaks simulation design described below and trends for three quantiles were fit simultaneously: 0.05, 0.1, and 0.15 using a $\lambda = n/5$. We use from one to four windows for each data size with an overlap of 500. The windows algorithm was run until the stopping criteria were met using $\epsilon_{abs} = 0.01$ and $\epsilon_{rel} = 0.001$. As is shown in \Fig{timing}, using 4 windows instead of one on data sizes of 55000 provides a factor of 3 decrease in computation time. The timing experiments were conducted on

>>>>

\documentclass[12pt]{article}
\RequirePackage{amsthm,amsmath,amsbsy,amsfonts}
\usepackage{bbm}
\usepackage{bm,graphicx,psfrag,epsf}
%\usepackage{enumerate}
\usepackage{natbib}
\usepackage{url} % not crucial - just used below for the URL 
\usepackage{placeins}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{subfig} %% For \subfloat command

@@

%% Math notation
\newcommand{\Tra}{^{\sf T}} % Transpose
\newcommand{\Inv}{^{-1}} % Inverse
\newcommand{\tr}{\operatorname{tr}} % Trace
\newcommand{\Ker}{\operatorname{Ker}} % Kernel
\def\vec{\mathop{\rm vec}\nolimits}
\newcommand{\amp}{\mathop{\:\:\,}\nolimits}
\newcommand{\bic}{\text{BIC}}
\newcommand{\VI}{\text{VI}}
\newcommand{\One}{\mathbbm{1}}

@@

	\bigskip
	\begin{abstract}
		We address the problem of estimating smoothly varying baseline trends in time series data using quantile trend filtering. We extend the basic framework to ensure non-crossing while estimating multiple trends simultaneously and implement an alternating direction method of moments (ADMM) algorithm for series that cannot be processed simultaneously. Timing experiments demonstrate the usefulness of our ADMM algorithm even in cases where the series can be processed simultaneously. We also address smoothing parameter selection and propose a modified criteria based on the extended Bayesian Information Criteria. We demonstrate both through simulation studies and an application to lost-cost air quality sensor data that our model provides better or comparable estimates than existing methods of non-parametric quantile trends and improves signal classification of low-cost air quality sensor output. 
	\end{abstract}

@@

	In applications spanning the fields of chemistry \citep{Ning2014}, macroeconomics \citep{yamada2017estimating}, environmental science \citep{brantley2014mobile}, and medical sciences \citep{pettersson2013algorithm, marandi2015qualitative}, scalar time series are observed and assumed to consist of a slowly varying trend and other more rapidly varying components. \cite{Kim2009} addressed the case when the series consisted of just a trend and a random noise component and proposed the use of \textit{$\ell_1$-trend filtering} to estimate trends that piecewise linear or piecewise polynomial. \cite{Tib2014} later demonstrated that empirically the trend filtering estimates adapt to the local level of smoothness better than the more common smoothing splines \Halley{Eric}{If he proved theorems, minimax stuff, we should state that.}. In the trend filtering problem, given observations $y(t)$ with $t=1,\ldots,n$, the trend, $\theta(t)$, is estimated by solving the following convex optimization problem. \Halley{Eric}{We maybe should define two objects $y(t)$ which is a function over the reals and $\V{y} \in \Real^n$, where $\VE{y}{i} = y(t_i)$.}
	\begin{eqnarray*}
	\underset{\V{\theta}}{\text{minimize}}\; \frac{1}{2} \lVert \V{y} - \V{\theta} \rVert_2^2 + \lambda \lVert \Mn{D}{k+1}\V{\theta} \rVert_1,
	\end{eqnarray*}
	where $\lambda \geq 0$ is a regularization parameter and the matrix $\Mn{D}{k+1} \in \Real^{(n - k -1) \times n}$ is the discrete difference operator of order $k+1$. To understand the purpose of penalizing $\Mn{D}{k+1}\V{\theta}$ consider the difference operator when $k = 0$.
	\begin{eqnarray*}
	\Mn{D}{1} & = & \begin{pmatrix}
	-1 & 1 & 0 & \cdots & 0 & 0 \\
	0 & -1 & 1 & \cdots & 0 & 0 \\
	\vdots & & & & & \\
	0 & 0 & 0 & \cdots & -1 & 1 \\
	\end{pmatrix}.
	\end{eqnarray*}
	Thus, $\lVert \Mn{D}{1}\V{\theta} \rVert_1 = \sum_{i=1}^{n-1} \lvert \theta_i - \theta_{i+1} \rvert$ which is just total variation denoising in one dimension. The penalty term incentivizes solutions which are piece-wise constant. For $k \geq 1$, the difference operator $\Mn{D}{k+1} \in \Real^{(n-k-1) \times n}$ is defined recursively as follows
	\begin{eqnarray*}
	\Mn{D}{k+1} & = & \Mn{D}{1}\Mn{D}{k}.
	\end{eqnarray*}
	By penalizing the $k+1$ fold composition of the discrete difference operator, we obtain solutions which are piecewise polynomials of order $k$.

	\Fig{tpn} shows that trend filtering provides excellent estimates of trends when the only components of the time series are the trend and random noise. In some applications, however, time series may also include, in addition to the trend and noise, a signal component that varies more rapidly than the trend. 
	\Fig{tpspn} shows that in these cases the trend filtering estimate over-estimates the trend in the places where signal is present.

	\begin{figure}[t]
		\centering
		\subfloat[Slow Trend + Noise]{\label{fig:tpn}
		\includegraphics[width = 0.45\linewidth]{Figures/trend_filter_eg1.png}}
		\subfloat[Slow Trend + Spikes + Noise]{\label{fig:tpspn}
		\includegraphics[width = 0.45\linewidth]{Figures/trend_filter_eg2.png}}
		\caption{Two examples of trend filtering solutions. The true trend is shown in black, while the estimated trend is shown in red. \Halley{Eric}{Could you explain to the reader what they should be seeing in the two panels, especially the one on the right? Don't leave to chance that a busy reviewer may miss the point.}}		
		\label{fig:trendfilter}
	\end{figure}

	One application in which the observed time series consists of a slowly varying baseline, non-negative signal, and rapidly varying noise is the output of lost-cost air quality sensors. The use of low-cost, portable, air quality sensors has increased dramatically in the last decade. These sensors can provide an un-calibrated measure of a variety of pollutants in near real time, but deriving meaningful information from sensor data remains a challenge \citep{snyder2013changing}. The ``SPod" is a low-cost sensor currently being investigated by researchers at the U.S. Environmental Protection Agency to detect volatile organic compound (VOC) emissions from industrial facilities \citep{thoma2016south}. To reduce cost and power consumption of the SPod, the temperature and relative humidity of the air presented to the photoionization detectors (PIDs) is not controlled and as a result the output signal exhibits a slowly varying baseline drift on the order of minutes to hours. \Fig{raw_spod} provides an example of 3 co-located SPod sensor outputs from near an industrial facility. While all of the sensors response to the pollutant signal, the amount of baseline drift varies from one node to another.

	\begin{figure}[b!]
		\includegraphics[width = \linewidth]{Figures/uncorrected_data.png}
		\caption{Example of 3 co-located SPod PID sensor readings.}		
		\label{fig:raw_spod}
	\end{figure}

	In this application, as well as those described in \cite{Ning2014}, \cite{marandi2015qualitative}, and \cite{pettersson2013algorithm}, it is not the mean trend that is desired but rather the baseline trend, which can be thought of as a trend in a low quantile of the data.  \cite{Koenker1978} were the first to propose substituting the squared loss typically used in regression with the check-loss function  \Eqn{check-loss} to estimate a conditional quantile instead of the conditional mean. 
	\begin{eqnarray}
	\label{eq:check-loss}
	 \rho_{\tau}(\V{z}) & = & \sum_{i=1}^n z_i(1-\One(z_i<0)),
	\end{eqnarray}
	where $\One(A)$ is 1 if its input $A$ is true and 0 otherwise.

	Later, \cite{KoenkerNgPortnoy1994} addressed the issue of estimated quantile curves by combining the check-loss penalty with trend filtering penalty with $k = 2$ which produces quantile trends that are piecewise linear. Other approaches to estimating quantile trends have also been proposed. Rather than using the trend filtering $\ell_1$-norm to penalize the discrete differences, \cite{nychka1995nonparametric} used the smoothing spline penalty based on the $\ell_2$-norm solving the optimization problem in \Eqn{smoothingspline}.

	\begin{equation*}
	\label{eq:smoothingspline}
	\rho_{\tau}(y - f) + \lambda\int (f''(t))^2 dt, 
 	\end{equation*}
where $f(t)$ is a smooth function of time and $\lambda$ is a tuning parameter that controls the degrees of smoothing. \Halley{Eric}{What is $\rho_{\tau}(\cdot)$ when its input is a function over the reals? It's defined above as a function of a vector input.} \cite{Oh2011} proposed an algorithm for solving this problem based on approximating the check-loss function with a similar differentiable function.

@@

	We propose to use the trend filtering penalty combined with the check-loss function to produce quantile trends that are piecewise quadratic. The formulation was proposed by \cite{Kim2009} as a possible extension of $\ell_1$-trend filtering but not studied. Moreover, we extend the basic framework to ensure non-crossing while modeling multiple quantiles. We also implement a parallel ADMM algorithm for series that are too large to be computed simultaneously and propose a modified criteria for choosing the smoothing parameter. We demonstrate through simulation studies that our proposed model provides better or comparable estimates of non-parametric quantile trends than existing methods and is a more effective method of drift removal for low-cost air quality sensors.

@@

	We combine the ideas of quantile regression and trend filtering. For a single quantile level $\tau$ the estimation of the quantile trend filtering model can be posed as the following convex optimization problem.
	\begin{eqnarray}
	\label{eq:quantile_trend}
	\underset{\V{\theta}}{\text{minimize}}\; \rho_\tau(\V{y} - \V{\theta}) + \lambda \lVert \Mn{D}{k+1} \V{\theta} \rVert_1,
	\end{eqnarray}
	where $\lambda$ is a non-negative regularization parameter. We address the problem of choosing $\lambda$ in Section \ref{sec:lambda_choice}. As with classic quantile regression, the quantile trend filtering problem is a linear program which can be solved by a number of free or commercial solvers. In many cases, including ours, we are interested in estimating multiple quantiles simultaneously. We also want to ensure that our quantile estimates are valid by enforcing the constraint that if $\tau_2 > \tau_1$ then $Q(\tau_2) \ge Q(\tau_1)$. \Halley{Eric}{Please define $Q$. The reader can infer what it is, but it's better not to put that on the reader.} Given quantiles $\{\tau_1, \ldots, \tau_J\}$ such that $\tau_1 < \tau_2 < \ldots < \tau_J$, the optimization problem becomes

	\begin{eqnarray}
	\label{eq:noncross_trend}
	\underset{\V{\theta}_1, \ldots, \V{\theta}_J}{\text{minimize}}\; \sum_{j=1}^J \left [\rho_{\tau_j}(\V{y} - \V{\theta}_{j}) + 
	\lambda_j \lVert \Mn{D}{k+1} \V{\theta}_j \rVert_1 \right ] \\
	 \text{subject to: }\; \theta_{1}(t) \le \theta_{2}(t) \le \ldots \le \theta_{J}(t) \text{ for all } t, \nonumber
	\end{eqnarray}
	where $\V{\theta}_j \in \Real^n$. The additional constraints are linear in the parameters so the non-crossing quantile trends can still be estimated by a number of available solvers. In the rest of this paper we rely on the commercial solver Gurobi \citep{gurobi} and its R package implementation, but it can easily be substituted for a free solver such as the Rglpk package by \cite{rglpk}. The number of parameters to be estimated in this problem is equal to the product of the number of observations $n$ and the number of quantiles of interest $J$. As the size of the data and the number of quantiles grows, all solvers will eventually break. \Halley{Eric}{Can we be more specific about what we mean when we say ``break?"}

@@

	\begin{align*}
	y_1(t) = y(t) & \mbox{~~if~~} 1 \le t \le u_{1}\\
	y_2(t) = y(t) & \mbox{~~if~~} l_{2} \le t \le u_{2} \\
	\vdots & \\
	y_M(t) = y(t) & \mbox{~~if~~} l_{M} \le t \le  n,
	\end{align*}
with boundaries $1 < l_{2} < u_{1} < l_{3} < u_{2} < l_{4} < u_{3} < \cdots< n$. An illustration is given in \Fig{windows}. Given quantiles $\tau_1 < \ldots < \tau_J$ to be estimated we define $\theta_{j,m}(t)$ as the value of the $\tau_j$\textsuperscript{th} quantile trend in window $M$ at time point $t$. In order to write out the constraint that the overlapping sections must be equal, we define a consensus variable
	 \begin{align*}
		 \bar{\theta}_{j,m} =&  g(\theta_{j, m-1}, \theta_{j,m}, \theta_{j,m+1}) \\
		 =& \begin{cases} 
			 \frac{\theta_{j,m-1}(t)+\theta_{j,m}(t)}{2} & \mbox{if~~} l_{m} \le t \le u_{m-1}  \\
			 \theta_{j,m}(t) & \mbox{if~~} u_{m-1} \le t \le l_{m+1}  \\
			 \frac{\theta_{j,m}(t)+\theta_{j,m+1}(t)}{2} & \mbox{if~~} l_{m+1} \le t \le u_{m}  
			 \end{cases},
	\end{align*}
	defining $\theta_{j,M+1} = \theta_{j,M}$ and $\theta_{j,0} = \theta_{j,1}$. Our windowed quantile trend optimization problem can then be written as 
	 \begin{align*}
		 \label{eq:quantile_windows}
		 &\underset{\theta_{1,m}, \ldots, \theta_{J,m}}{\min}\; \sum_{m=1}^M\sum_{j=1}^J \left [\rho_{\tau_j}(y_m - \theta_{j,m}) + 
		 \lambda_j \lVert \Mn{D}{k+1} \theta_{j,m} \rVert_1 \right ] \\
		 &\text{subject to: }\; \\
		 &\qquad \theta_{1,m}(t) \le \theta_{2,m}(t) \le \ldots \le \theta_{J,m}(t) \text{ for all } m,t \\
		 &\qquad \theta_{j,m}(t) = \bar{\theta}_{j,m}(t) \text{ for all } j, m, t
	 \end{align*}
	 Defining Lagrange multipliers $\omega_{j,m}$, and penalty parameter $\gamma > 0$ we can write the augmented Lagrangian for finding the trends in window $m$:
	 \begin{equation*}
	 \mathcal{L}(\theta_{j,m}, \bar{\theta}_{j,m}, \omega_{j,m}) = \sum_{j=1}^J\rho_{\tau_j}(y_m - \theta_{j,m})+\lambda \lVert \Mn{D}{k+1}\theta_{j,m}\rVert_1 +  \omega_{j,m}\Tra(\theta_{j,m} - \bar{\theta}_{j,m}) + 
	 \frac{\gamma}{2}||\theta_{j,m} - \bar{\theta}_{j,m}||_2^2
	 \end{equation*}
	 We then estimate the trend separately in each window, which can be done in parallel, while constraining the overlapping pieces of the trends to be equal as outlined in algorithm 1.

@@

	\begin{algorithm}
		\caption{ADMM algorithm for quantile trend filtering with windows}\label{euclid}
		\begin{algorithmic}		
			\State Define $\M{D} = \Mn{D}{k+1}$. 
			\State \textbf{initialize:} \\
				 $\Vn{\theta}{0}_{j,m} \gets \arg \min \sum_{j=1}^J\rho_{\tau_j}(\V{y}_m - \V{\theta}_{j,m})+\lambda \lVert \M{D}\V{\theta}_{j,m}\rVert_1$ subject to $\theta_{1,m}(t) < \ldots<\theta_{J,m}(t)$ for all $t$. \\		
			 	 $\Vn{\omega}{0}_{j,m} \gets \V{0}$	
			\Repeat{}
			\State  		
			$\bar{\theta}_{j,m}^{(q)} \gets g(\theta_{j, m-1}^{(q-1)}, \theta_{j,m}^{(q-1)}, \theta_{j,m+1}^{(q-1)})$
			\State 
			$\omega_{j,m}^{(q)} \gets \omega_{j,m}^{(q-1)} + \gamma(\theta_{j,m}^{(q-1)} - \bar{\theta}_{j,m}^{(q)})$	
			\State
				$\theta_{j,m}^{(q)} \gets \arg\min \mathcal{L}(\theta_{j,m}, \bar{\theta}_{j,m}^{(q-1)}, \omega_{j,m}^{(q-1)})$			
			 subject to $\theta_{1,m}(t) < \ldots<\theta_{J,m}(t)$ for all $t$.
			\Until {convergence}
			\State \textbf{return} Non-overlapping sequence of $\bar{\theta}_{j,m}^{(q)}$ for all $j$, $m$.

@@

	We measure convergence use the stopping criteria described by \cite{boyd2011distributed}. The criteria are based on the primal and dual residuals which represent the residuals for the primal and dual feasibility, respectively. The primal residual, 
	\begin{eqnarray*}
	r_p^{(q)} & = & \sqrt{\sum_{m=1}^M\sum_{j=1}^J\lVert\theta_{j,m}^{(q)} - \bar{\theta}_{j,m}^{(q)}\rVert_2^2},
	\end{eqnarray*}
	represents the difference between the trend values in the windows and the consensus trend value while the dual residual,
	\begin{eqnarray*}
	r_d^{(q)} & = & \gamma\sqrt{\sum_{m=1}^M \sum_{j=1}^J\lVert\bar{\theta}_{j,m}^{(q)} - \bar{\theta}_{j,m}^{(q-1)}\rVert_2^2},
	\end{eqnarray*}
	represents the change in the consensus variable from one iterate to the next. The algorithm is stopped when 
	\begin{eqnarray*}
		r_p^{(q)} & < &\epsilon_{\text{abs}}\sqrt{nJ} + \epsilon_{\text{rel}}\underset{m}{\max}\left[\max 
		\left(\sqrt{\sum_{j=1}^J \lVert\theta_{j,m}^{(q)}\rVert_2^2}, \sqrt{\sum_{j=1}^J \lVert \bar{\theta}_{j,m}^{(q)} \rVert_2^2} \right )\right]\\
		r_d^{(q)} & < & \epsilon_{\text{abs}}\sqrt{nJ} + \epsilon_{\text{rel}}\sqrt{\sum_{m=1}^M\sum_{j=1}^J\lVert \omega_{j,m}^{(q)}\rVert_2^2}
	\end{eqnarray*}

	\begin{figure}
		\centering
		\includegraphics[width = 0.8\linewidth]{Figures/admm_windows.png}
		\caption{Trend fit with our ADMM algorithm with 3 windows which converged in 7 iterations compared to trend from simultaneous fit.}
	\end{figure}

	Timing experiments illustrate the advantages of using our ADMM algorithm even on datasets where solving the problem simultaneously is possible. For each data size, $n$, 25 datasets were simulated using the peaks simulation design described below and trends for three quantiles were fit simultaneously: 0.05, 0.1, and 0.15 using a $\lambda = n/5$. We use from one to four windows for each data size with an overlap of 500. The windows algorithm was run until the stopping criteria were met using $\epsilon_{\text{abs}} = 0.01$ and $\epsilon_{\text{rel}} = 0.001$. \Fig{timing} shows that using 4 windows instead of one on data sizes of 55000 provides a factor of 3 decrease in computation time. The timing experiments were conducted on
